\huge Gap formation and stability in non-isothermal protoplanetary disks \large [2014 CITA Summer Student Program]

### 2014-07-24 poster/poster.tex
poster completed

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            \begin{block}{{\Large Introduction}}
              \justifying
                  {\large              
                    {\bf
                      The formation of large-scale vortices at planetary gap edges has been
                      suggested to explain observations of transition
                      disks with lopsided dust
                      distributions. Thus far, theoretical disk-planet
                      models have employed locally isothermal disks, while 
                      the linear theory of the vortex-formation 
                      instability was originally developed for adiabataic disks. We
                      generalize the study of planetary gap edge
                      instabilities to non-isothermal disks by
                      including an energy equation with variable
                      cooling rate, and study its effect on the
                      formation and evolution of gap-edge vortices. 
                    }
                  }
            \end{block}
            \vfill

            \begin{block}{{\Large Disk-planet model with simple
                  cooling}}
              \justifying
             We numerically evolve the 2D fluid equations including a
             gap-opening giant planet, treated as an external
             potential on a fixed circular orbit. We include in the
             energy equation  
        a simple source term in the form 
	\begin{align*}
	\frac{\partial e}{\partial t} = \cdots + \frac{e-e_{t=0}}{t_\mathrm{cool}},
	\end{align*}
        and $t_\mathrm{cool} = \beta \Omega_k^{-1}$, where $\Omega_k$
        is the Keplerian frequency. The energy density
        $e=p/(\gamma-1)$ where $p$ is the pressure field and
        $\gamma=1.4$ is the adiabatic index. We choose $\beta$ indirectly 
        through $\tilde{\beta}$ such that
        \begin{align*}
        t_\mathrm{cool, gap\,edge} = \tilde{\beta} t_\mathrm{lib, gap\,edge},
        \end{align*}
\vspace{-1.6cm}
        \begin{itemize}
        \item  $t_\mathrm{lib, gap\,edge}=2\pi/|\Omega_k(r_p)-\Omega_k(r_\mathrm{edge})|$ is the time inbetween
          successive encounters of a fluid element at the gap
          edge and the planet's azimuth
        \item $\tilde{\beta}\ll 1 \Rightarrow$ fluid at the gap edge
          returns to its initial temperature inbetween going through planet-induced shocks 
        \end{itemize}
        \end{block}
        \vfill

@@

                    \item Slowering cooling rate increases

@@

            \begin{block}{\Large{Non-linear evolution of gap edge
                  vortices}}
              \justifying
%                  {\bf `lifetime' plot of m=1 for diff beta, snapshots
%                    (before, during, after vortex decay). quote rossby
%                  number when vortices in quasi-steady state}
              We include the planet throughout. Generic evolution independent of $t_\mathrm{cool}$:
              \begin{itemize}
              \item Single vortex in quasi-steady state from vortex merging
              \item Continous vorticity source from disk-planet interaction
                $\to$ vortex intensifies
              \item Vortex dissipates quickly after its Rossby number
                $Ro$ minimizes, due to vortex-induced shocks and/or smoothing out the
                gap edge  
              \end{itemize}
                  \begin{figure}
                    \centering
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_Before}
                      \caption{$Ro=-0.15$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_During}
                      \caption{$Ro=-0.43$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_After}
                      \caption{$Ro=-0.05$}
                    \end{minipage}
                    \hfill
                  \end{figure}

>>>>

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            \begin{block}{{\Large Introduction}}
              \justifying
                  {\large              
                    {\bf
                      Large-scale vortex formation at planetary gap edges has been
                      suggested to explain observations of transition
                      disks with lopsided dust
                      distributions. Thus far, theoretical
                      models have employed locally isothermal disks, while 
                      the linear theory of the vortex-formation 
                      instability was originally developed for adiabatic disks. We
                      generalize the study of planetary gap edge
                      instabilities to non-isothermal disks by
                      including an energy equation with variable
                      cooling rate, and study its effect on the
                      formation and evolution of gap-edge vortices. 
                    }
                  }
            \end{block}
            \vfill

            \begin{block}{{\Large Disk-planet model with simple
                  cooling}}
              \justifying
             We numerically evolve the 2D fluid equations including a
             gap-opening giant planet, treated as an external
             potential on a fixed circular orbit. We include in the
             energy equation  
        a simple source term in the form 
	\begin{align*}
	\frac{\partial e}{\partial t} = \cdots + \frac{e-e_{t=0}}{t_\mathrm{cool}},
	\end{align*}
        and $t_\mathrm{cool} = \beta \Omega_k^{-1}$, where $\Omega_k$
        is the Keplerian frequency. The energy density
        $e=p/(\gamma-1)$ where $p$ is the pressure field and
        $\gamma=1.4$ is the adiabatic index. We choose $\beta$ indirectly 
        through $\tilde{\beta}$ such that
        \begin{align*}
        t_\mathrm{cool, gap\,edge} = \tilde{\beta} t_\mathrm{lib, gap\,edge},
        \end{align*}
\vspace{-1.6cm}
        \begin{itemize}
        \item  $t_\mathrm{lib, gap\,edge}=2\pi/|\Omega_k(r_p)-\Omega_k(r_\mathrm{edge})|$ is the time in between
          successive encounters of a fluid element at the gap
          edge and the planet's azimuth
        \item $\tilde{\beta}\ll 1 \Rightarrow$ fluid at the gap edge
          returns to its initial temperature in between going through planet-induced shocks 
        \end{itemize}
        \end{block}
        \vfill

@@

                    \item Slower cooling rate increases

@@

            \begin{block}{\Large{Non-linear evolution of gap edge
                  vortices}}
              \justifying
%                  {\bf `lifetime' plot of m=1 for diff beta, snapshots
%                    (before, during, after vortex decay). quote rossby
%                  number when vortices in quasi-steady state}
              We include the planet throughout. Generic evolution independent of $t_\mathrm{cool}$:
              \begin{itemize}
              \item Single vortex in quasi-steady state from vortex merging
              \item Continuous vorticity source from disk-planet interaction
                $\to$ vortex intensifies
              \item Vortex dissipates quickly after its Rossby number
                $Ro$ minimizes, due to vortex-induced shocks and/or smoothing out the
                gap edge  
              \end{itemize}
                  \begin{figure}
                    \centering
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_Before}
                      \caption{$Ro=-0.15$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_During}
                      \caption{$Ro=-0.43$}
                    \end{minipage}
                    \hfill
                    \begin{minipage}{0.3\textwidth}
                      \includegraphics[width=\textwidth]{Posterfig_After}
                      \caption{$Ro=-0.05$}
                    \end{minipage}
                    \hfill
                  \end{figure}
